---
title: "SDS315 - HW 3"
author: "Aggie Angeles"
date: "2024-01-31"
output: html_document

---

```{r setup, include=FALSE}

# Libraries 
library(ggplot2)
library(tidyverse)
library(kableExtra)
library(mosaic)
library(dplyr)

# DATA SET 
creat = read.csv("creatinine.csv")
market = read.csv("marketmodel.csv")
covid = read.csv("covid.csv")
milk = read.csv("milk.csv")



```

## *Problem 1*

### *Creatinine*

```{r echo = FALSE, message = FALSE}

# Scatterplot +linear regress.
ggplot(creat) +
  geom_point(aes(x = age, y = creatclear)) +
  labs(title = " Age v.s Creatine Clearance Rate", 
       x = "Age", y = "Creatine Clearance Rate (mL/minute)") +
  geom_smooth(aes(x=age, y=creatclear), method='lm')
```

```{r echo=FALSE, include=FALSE}
expect_creat = lm(creatclear ~ age, data = creat)
coef(expect_creat)

148 - .6*(55)

```




Given our equation based off the regression model for the relationship for creatinine clearance rate based off age, we should expect a 55 year-old to have a creatinine clearance rate of 115 mL/minute.By taking the data frame and running it through a lm function and then using the coef function, we are able to calculate and predict a corresponding value.


Based off the equation, the increase of a person's age results in a decrease of .6 mL/minute per year in a person's creatinine clearance rate.
When reviewing the basic components of the function, we understand that 148 is the baseline creatinine clearance rate, and the coefficient -.6 is directly affecting the inputted age.
It is therefore the slope, given y = mx + b, of the linear regression model.
The negative sign before the .6 indicates that it causes a decrease in the value while the .6 indicates how much the creatinine clearance rate is changing by.

All comparsions are made using our defined linear regression model.
The expected creatinine for a 40 year-old is a rate of 124 mL/minute versus a rate of 135, which is higher by 11 mL/minute.
The actual is lower than what is expected of a person of their age.
The expected creatinine for a 60 year-old is a rate of 112 mL/minute, which exactly the actual rate proposed in the question.
To visualize this I have written out the equation as follows: By taking the equation y = 148 - .6(age), and inputting 40 as our age we output y = 124 mL/minute.
Then taking that value and subtracting it by 135 mL/minute we get a difference of 11 mL/minute.
Likewise we input 60 for the age and a return a y-value of 112 mL/minute, where the expected value equals the actual value in mL/minute.



```{r echo=FALSE, include=FALSE}
# Expected 40yrs
148 - .6*(40)
# = 124
# actual is 135

# Expected 60yrs
148 - .6*(60) 
# = 112
# actual is 112

```

## *Problem 2*

### *Capital Asset Pricing Model (CAPM).*

```{r echo = FALSE, message = FALSE}

ggplot(market) +
  geom_point(aes(x = SPY, y = AAPL)) + labs(title = "Distribution of Apple Stock on the Return of S&P Stocks", x = "S&P", y = "Apple")

ggplot(market) +
  geom_point(aes(x = SPY, y = GOOG)) + labs(title = "Distribution of Google Stock on the Return of S&P Stocks", x = "S&P", y = "Google")

ggplot(market) +
  geom_point(aes(x = SPY, y = MRK)) + labs(title = "Distribution of Merek Stock on the Return of S&P Stocks", x = "S&P", y = "Merek")

ggplot(market) +
  geom_point(aes(x = SPY, y = JNJ)) + labs(title = "Distribution of Johnson & Johnson Stock on the Return of S&P Stocks", x = "S&P", y = "Johnson & Johnson")

ggplot(market) +
  geom_point(aes(x = SPY, y = WMT)) + labs(title = "Distribution of  Wal-Mart Stock on the Return of S&P Stocks", x = "S&P", y = "Wal-Mart")

ggplot(market) +
  geom_point(aes(x = SPY, y = TGT)) + labs(title = "Distribution of  Target Stock on the Return of S&P Stocks", x = "S&P", y = "Target" )


```

```{r echo=FALSE, include=FALSE}

Model1 = lm(SPY ~ AAPL, data=market)
coef(Model1)
rsquared(Model1)

Model2 = lm(SPY ~ GOOG, data=market)
coef(Model2)
rsquared(Model2)

Model3 = lm(SPY ~ MRK, data=market)
coef(Model3)
rsquared(Model3)

Model4 = lm(SPY ~ JNJ, data=market)
coef(Model4)
rsquared(Model4)

Model5 = lm(SPY ~ WMT, data=market)
coef(Model5)
rsquared(Model5)

Model6 = lm(SPY ~ TGT, data=market)
coef(Model6)
rsquared(Model6)
```

```{r echo = FALSE}


StocksTable <- tibble(Ticker = c("AAPL", "GOOG", "MRK", "JNJ", "WMT", "TGT"), 
                      Intercept = c(.00076, .00016,.00056, .00046, .00026, .00011),
                      Slope = c(.01254, .6504, .67789, .74121, .54978, .35029),
                      R2 = c(.01336, .64830, .48370, .50194, .28532, .24788))




StockKbl = StocksTable %>% 
  kbl(col.names = c("Ticker", "Intercept", "Slope", "R2"), caption = "Capital Asset Pricing Model Regression Data")

StockKbl %>% 
  kable_classic(full_width = F, html_font = "Cambria", font_size = 12)


```


Beta is a measure of systematic risk. It is a correlation between a given stock and the overall market. A stock's return in percentage change is indicative of a one percentage in a market portfolio.The smaller the beta, the less systematic risk there is. Likewise, the greater the beta, the higher systematic risk there is. In particular, a beta less than 1 is classified as low systematic risk whereas a beta greater than 1 is said to have high systematic risk. When a beta of a stock is 0 it has no systematic risk because it does not react to changes in the market. Beta values also make logical sense in that when a market is bad, stocks or goods that are most effected are those that are typically nonessential. That said, it would make sense for Walmart or Target, which are grocery stores, to have a low beta value in comparison to Google. 
Apple has the lowest systematic risk as indicated by its beta, or R2 value, which is .01336. 
This means that with respect to the market as a whole, Apple is less volatile to changes both positive and negative. For every one percent change in the market, Apple shifts by .01336, in either the positive or negative direction in direct correlation to the market's sign.
In that same line of thinking, Google has the highest systematic risk with its beta of .64830. 

The table details the data of six different stocks: Apple (AAPL), Google (GOOG), Merck (MRK), Johnson and Johnson (JNJ), Wal-Mart (WMT), and Target (TGT). We took the linear regression of these stocks on the return of the S&P 500 and found its intercept, slope, and beta (R2).




## *Problem 3*

### *COVID-19 Deaths: Italy and Spain*

```{r echo = FALSE, include=FALSE}
# ITALY
italyCovid <- covid %>% 
  filter(country == "Italy")

# linear model for log(death) - ITALY
lmItaly = lm(log(deaths) ~ days_since_first_death, data = italyCovid)

coef(lmItaly)


# SPAIN
spainCovid <- covid %>% 
  filter(country == "Spain")

# linear model for log(death) - ITALY
lmSpain = lm(log(deaths) ~ days_since_first_death, data = spainCovid)

coef(lmSpain)

```

The estimated growth rate of Covid death cases during this time was 18.3% with a doubling time of about 4 days. 
The estimated growth rate of Covid death cases during this time was 27.6% with a doubling time of about 3 days. 



```{r echo = FALSE, mesage = FALSE}
ggplot(covid) +
  geom_point(aes(x = days_since_first_death, y = log(deaths), col = country)) + labs(title = "Total Covid Death Cases in Italy & Spain Over Time ", x = "Days Since First Death", y = "Covid Deaths") + theme(text=element_text(size=12,  family="serif")) +
  geom_smooth(aes(x= days_since_first_death, y=log(deaths), col = country), method='lm')
```



## *Problem 4*

### *Milk: Elasticity*

```{r echo = FALSE, message = FALSE}

# Milk Elasticity: log base values
ggplot(milk) +
   geom_point(aes(x=log(price), y=log(sales))) +
  geom_smooth(aes(x=log(price), y=log(sales)), method='lm') +
  labs(title = "Milk Elasticity Data", x = "Price", y = "Sales") + theme(text=element_text(size=12,  family="serif"))


```

```{r echo = FALSE, include=FALSE}
# Slope & Intercept Regression
lmMilk = lm(log(sales) ~ log(price), data=milk)
coef(lmMilk)

```

With this specific data frame, we took the log base values of both the x & y. Using first the lm() function and then from there inputting that information into the coef() to get out slope and y-intercept points. The y-intercept was 4.72 sales with a decreasing slope of about $1.62 per sale. Translating this in terms our power law, the equation is 112.2 multiplied by x to the power of -1.62.

The estimated price elasticity of the demand of milk is -1.62, meaning that when the price of milk increases by one percent, people are less willingly to want to buy milk by 1.62% on average.




